---
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    x = paste(x, collapse = '.')
    if (!grepl('\\.svg', x)) return(hook_plot(x, options))
    # read the content of the svg image and write it out without <?xml ... ?>
    paste0('<img src= "./', x, '">')
  })
})
```

# tvdiff

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/natbprice/tvdiff.svg?branch=master)](https://travis-ci.org/natbprice/tvdiff)

The **tvdiff** package is an R translation of the Matlab implementation of the 
Total Variation Regularized Numerical Differentiation algorithm by Rick Chartrand.
The package implements the methods found in Rick Chartrand,"Numerical 
differentiation of noisy, nonsmooth data," ISRN Applied Mathematics, Vol. 2011,
Article ID 164564, 2011.

## Installation

The **tvdiff** package is currently only available from Github.

```{r, eval = FALSE}
devtools::install_github("natbprice/tvdiff")
```

## Example
A simple example based on the function $f(x) = \mid x - 0.5 \mid$ with 
Gaussian noise of standard deviation 0.05. The derivative is estimated from the 
noisy observations using Total Variation Regularized Differentiation. A
prediction of the original function is obtained from the estimated derivative 
through numerical integration.

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load packages
library(tvdiff)
library(tidyr)
library(dplyr)
library(ggplot2)
```

### Load demo data
```{r, echo = TRUE, message = FALSE, warning = FALSE}
data("smalldemodata")
str(smalldemodata)
```

### Use `TVRegDiffR` to perform numerical integration on the demo data
```{r, echo = TRUE, message = FALSE, warning = FALSE, dev = 'svg', fig.height=3, fig.width=6}
# Build dataframe
smallEx <-
  smalldemodata %>%
  # Estimate derivative
  mutate(dxdt =
           TVRegDiffR(
             data = obs,
             iter = 1e3,
             alph = 0.2,
             scale = "small",
             ep = 1e-6,
             dx = 0.01
           )[-1]) %>% 
  # Simple numerical integration
  mutate(dx = lead(x) - x,
         pred = obs[1] + cumsum(dxdt*dx),
         true = abs(x - 0.5)) %>% 
  # Collect in long form
  gather(key, value, -x, -dx)

# Plot observed vs predicted
ggplot(smallEx %>% 
         filter(key != "dxdt"), 
       aes(x = x, y = value, color = key, linetype = key)) +
  geom_line() +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  labs(x = "x",
       y = "f(x)")

# Plot derivative
ggplot(smallEx %>% 
         filter(key == "dxdt"), 
       aes(x = x, y = value)) +
  geom_point() +
  theme_minimal() +
  labs(x = "x",
       y = "f'(x)")
```


## References
Rick Chartrand, "Numerical differentiation of noisy, nonsmooth data," ISRN Applied Mathematics, Vol. 2011, Article ID 164564, 2011.

https://sites.google.com/site/dnartrahckcir/home/tvdiff-code

Python translation by Simone Sturniolo:
https://github.com/stur86/tvregdiff
