---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    x = paste(x, collapse = '.')
    if (!grepl('\\.svg', x)) return(hook_plot(x, options))
    # read the content of the svg image and write it out without <?xml ... ?>
    paste0('<img src= "./', x, '">')
  })
})
```
# tvdiff

The **tvdiff** package is a simple R wrapper for the Python translation of the 
original Matlab code by Rick Chartrand. The package implements the methods found in 
Rick Chartrand,"Numerical differentiation of noisy, nonsmooth data," ISRN 
Applied Mathematics, Vol. 2011, Article ID 164564, 2011.

## Installation

The **tvdiff** package is currently only available from Github.The package 
depends on reticulate and requires Python to be installed.

```{r, eval = FALSE}
devtools::install_github("natbprice/tvdiff")
```

## Example
A simple example based on the function $f(x) = \mid x - 0.5 \mid $ with 
Gaussian noise of standard deviation 0.05.

```{r, echo = FALSE, message = FALSE, warning = FALSE, dev = 'svg', fig.height=3, fig.width=6}
library(tvdiff)
library(tidyr)
library(dplyr)
library(ggplot2)

# Load demo data
data("smalldemodata")

# Build dataframe
smallEx <-
  smalldemodata %>%
  # Estimate derivative
  mutate(dxdt =
           TVRegDiffR(
             data = obs,
             iter = 1e3,
             alph = 0.2,
             scale = "small",
             ep = 1e-6,
             dx = 0.01
           )[-1]) %>% 
  # Simple numerical integration
  mutate(dx = lead(x) - x,
         pred = obs[1] + cumsum(dxdt*dx),
         true = abs(x - 0.5)) %>% 
  # Collect in long form
  gather(key, value, -x, -dx)

# Plot observed vs predicted
ggplot(smallEx %>% 
         filter(key != "dxdt"), 
       aes(x = x, y = value, color = key, linetype = key)) +
  geom_line() +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  labs(x = "x",
       y = "f(x)")

# Plot derivative
ggplot(smallEx %>% 
         filter(key == "dxdt"), 
       aes(x = x, y = value)) +
  geom_point() +
  theme_minimal() +
  labs(x = "x",
       y = "f'(x)")
```


## References
Rick Chartrand, "Numerical differentiation of noisy, nonsmooth data," ISRN Applied Mathematics, Vol. 2011, Article ID 164564, 2011.

https://sites.google.com/site/dnartrahckcir/home/tvdiff-code

R package based on a Python translation by Simone Sturniolo:
https://github.com/stur86/tvregdiff
