---
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
library(knitr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

local({
  hook_plot = knit_hooks$get('plot')
  knit_hooks$set(plot = function(x, options) {
    x = paste(x, collapse = '.')
    if (!grepl('\\.svg', x)) return(hook_plot(x, options))
    # read the content of the svg image and write it out without <?xml ... ?>
    paste0('<img src= "./', x, '">')
  })
})
```

# tvdiff

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/natbprice/tvdiff.svg?branch=master)](https://travis-ci.org/natbprice/tvdiff)
[![Coverage status](https://codecov.io/gh/natbprice/tvdiff/branch/master/graph/badge.svg)](https://codecov.io/github/natbprice/tvdiff?branch=master)

The **tvdiff** package is a solution to the problem of how to estimate the rate of change (i.e., derivative) when available data is noisy and nonsmooth. If the data is noisy, then simple finite difference methods of estimating the derivative are often inaccurate. If the data generating process is nonsmooth, then fitting a smooth model to the data in order to estimate the derivative may yield inaccurate estimates.

This package is based on the Matlab implementation of the Total Variation Regularized Numerical Differentiation algorithm by Rick Chartrand. This package uses a C++ method for the preconditioned conjugate method based on the cPCG package. See [references](#references) section.

## Installation

The **tvdiff** package is currently only available from Github.
```{r install_package, eval = FALSE}
# Install development version from GitHub (install function requires devtools package)
devtools::install_github("natbprice/tvdiff")
```

__Installation and use requirements__:
- R v. 4.0.1+
- Fortran
- _For mac OS users_: see [instructions](https://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/) regarding R compiler tools for R package **Rcpp**


## Example
A simple example based on the function $f(x) = \mid x - 0.5 \mid$ with Gaussian noise of standard deviation 0.05. The derivative is estimated from the noisy observations using Total Variation Regularized Differentiation. A prediction of the original function is obtained from the estimated derivative through numerical integration.

```{r load_package, echo = FALSE, message = FALSE, warning = FALSE}
# load packages
library(tvdiff)
```

### Load demo data
Example data is included in the package for demonstration.

```{r load_data, echo = TRUE, message = FALSE, warning = FALSE}
data("smalldemodata")
str(smalldemodata)
```

### Estimate derivative
The `TVRegDiffR` function is used to estimate the derivative. The main tuning parameter is the regularization parameter `alph`. Start by varying `alph` by 
orders of magnitude until reasonable results are obtained. The number of iterations `iter` must also be specified since there is no default stopping criteria.

```{r estimate_derivative, echo = TRUE, message = FALSE, warning = FALSE}

# Unpack data
x <- smalldemodata$x
obs <- smalldemodata$obs
true <- smalldemodata$true

# True derivative
dydx_true <- rep(-1, length(x))
dydx_true[x > 0.5] <- 1
dx <- x[2] - x[1]

# Extimate derivative
dydx <- TVRegDiffR(
  data = obs, # Vector of data to be differentiated
  iter = 1e3, # Number of iterations
  alph = 0.2, # Regularlization parameter
  dx = dx     # Grid spacing 
  )
dydx <- dydx[-1]

# Prediction
pred <- obs[1] + cumsum(dydx*dx)
```

### Plot results
For this simple example we can compare the estimates to the true values.

```{r plot_results, echo = FALSE, dev = 'svg', fig.height=4, fig.width=6}
# Plot observed vs predicted
plot(x, obs, pch = 20, col = "#57575F", ylab = "y")
lines(x, pred, col = "#CA3542", lty = 2, lwd = 2)
lines(x, true, col = "#27647B", lty = 1, lwd = 2)
legend("bottomleft", 
       legend = c("Observed", "Predicted", "True"),
       col = c("#57575F", "#CA3542", "#27647B"),
       pch = c(20, NA, NA),
       lty = c(NA, 2, 1),
       cex = 0.8,
       lwd = c(NA, 2, 2))

# Plot derivative
plot(x, dydx, ylim = c(-1,1), col = "#CA3542", type = "l", lty = 2, lwd = 2,
     ylab = expression(dy/dx))
lines(x, dydx_true, col = "#27647B", lty = 1, lwd = 2)
legend("topleft", 
       legend = c("Predicted", "True"),
       col = c("#CA3542", "#27647B"),
       lty = c(2, 1),
       lwd = c(2, 2),
       cex = 0.8)
```


## References {#references}
Rick Chartrand, “Numerical Differentiation of Noisy, Nonsmooth Data,”
ISRN Applied Mathematics, vol. 2011, Article ID 164564, 11 pages, 2011.
https://doi.org/10.5402/2011/164564.

**Matlab code:**
https://sites.google.com/site/dnartrahckcir/home/tvdiff-code

**Python translation:**
https://github.com/stur86/tvregdiff

**C++ implementation of preconditioned conjugate gradient method:**
https://github.com/styvon/cPCG
