---
title: "Example"
author: "Nathaniel Price"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(tvdiff)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Toy Problem: Tuning regularization parameter
This example is based on the article "Numerical Differentiation of
Noisy, Nonsmooth Data" by Rick Chartrand (http://dx.doi.org/10.5402/2011/164564).
We use the function $f(x) = \mid x - 0.5 \mid$ with Gaussian noise of standard deviation 0.05. First, we load and unpack the example data.

```{r load_data, warning = FALSE, fig.height=4, fig.width=6}
# Load small demo data 
data("smalldemodata")

# Unpack data
x <- smalldemodata$x
obs <- smalldemodata$obs
true <- smalldemodata$true

# True derivative
dydx_true <- rep(-1, length(x))
dydx_true[x > 0.5] <- 1
dx <- x[2] - x[1]

```

Next, the `TVRegDiffR` function is used to estimate the derivative. The main tuning parameter is the regularization parameter `alph`. We start by varying `alph` by 
orders of magnitude.

```{r tuning}
# Vector of regularization parameters to try
alph <- c(1e-3, 1e-1, 1)

# Preallocate space for results
dydx <- matrix(ncol = length(alph), nrow = length(x))
pred <- matrix(ncol = length(alph), nrow = length(x))

# Estimate derivative using different regularization parameters
for(i in seq_along(alph)) {
  # Extimate derivative
  estimate <- TVRegDiffR(
    data = obs, # Vector of data to be differentiated
    iter = 1e3, # Number of iterations
    alph = alph[i], # Regularlization parameter
    dx = dx     # Grid spacing 
    )
  dydx[, i] <- estimate[-1]
  
  # Prediction
  pred[, i] <- obs[1] + cumsum(dydx[, i]*dx)
}
```

We plot the results to determine if the regularization parameter is reasonable.

```{r smalldemo_plot1, fig.height=4, fig.width=6}
# Line style and color for plot
lty <- c(2, 2, 2)
col <- c("#1b9e77", "#7570b3", "#e7298a")

# Initialize plot with observed and true
plot(x, obs, pch = 20, col = "#57575F", ylab = "y")
lines(x, true, col = "#d95f02", lty = 1, lwd = 2)

# Add lines for each prediction
for(i in seq_along(alph)) {
  lines(x, pred[, i], col = col[i], lty = lty[i], lwd = 2)
}

# Add legend to plot
legend("bottomleft", 
       legend = c("Observed", "True", paste0("alph = ", alph)),
       col = c("#57575F", "#d95f02", col),
       pch = c(20, rep(NA, length(alph) + 1)),
       lty = c(NA, 1, lty),
       lwd = c(NA, rep(2, length(alph) + 1)),
       cex = 0.8)

```

```{r, fig.height=4, fig.width=6}
# Line style and color for plot
lty <- c(2, 2, 2)
col <- c("#1b9e77", "#7570b3", "#e7298a")

# Initialize plot with observed and true
plot(x, dydx_true, col = "#d95f02", ylim = c(-1.5,1.5), 
     ylab = expression(dy/dx), type = "l")

# Add lines for each prediction
for(i in seq_along(alph)) {
  lines(x, dydx[, i], col = col[i], lty = lty[i], lwd = 2)
}

# Add legend to plot
legend("topleft", 
       legend = c("True", paste0("alph = ", alph)),
       col = c("#d95f02", col),
       lty = c(1, lty),
       lwd = 2,
       cex = 0.8)
```

When using a value of `alph = 1e-3` the estimated derivative is too wiggly and therefore the regularization parameter is too low. Using a value of `alph = 1` the prediction does not match the observed data and therefore the regularization parameter is too high. Using a regularization parameter of `alph = 0.1` yields reasonable results.


## Application to real data: Respirometry example 
This example is based on the article "Numerical Differentiation of
Noisy, Nonsmooth Data" by Rick Chartrand (http://dx.doi.org/10.5402/2011/164564). However, since the tvdiff package does not implement the large scale algorithm we use only a subset of data and apply the small scale algorithm. The results are similar to those obtained in the original article.

```{r largeDemo, fig.height=4, fig.width=6}
# Load data
data("largedemodata")

# Subsample data in order to use 'small' algorithm
ind <- seq(1, nrow(largedemodata), 1e2)
largedemodata <- largedemodata[ind, ]

# Unpack data
x <- largedemodata$x
obs <- largedemodata$obs
dx <- x[2] - x[1]

# Extimate derivative
dydx <- TVRegDiffR(
  data = obs,
  iter = 40,
  alph = 1e4,
  scale = "small",
  ep = 1e-6,
  dx = dx
  )
dydx <- dydx[-1]

# Prediction
pred <- obs[1] + cumsum(dydx*dx)

# Plot observed vs predicted
plot(x, obs, pch = 20, col = "#57575F", xlab = "x", ylab = "y")
lines(x, pred, col = "#CA3542", lty = 2, lwd = 2)
legend("topleft", 
       legend = c("Predicted"),
       col = c("#CA3542"),
       lty = 2:1,
       cex = 0.8,
       lwd = 2)

# Plot derivative
plot(x, dydx, type = "l", col = "#CA3542", ylab = expression(dy/dx))

```

## A more challenging toy problem
We create a more challenging example based on the piecewise function $f(x) = (x+2)^2$ when $x<0$ and $f(x) = (x-2)^2$ when $x>0$ with Gaussian noise with a standard deviation of 0.5. There is considerable discrepancy between the estimated derivative and the true derivative. However, the results are much better than a 
simple finite difference method.

```{r toy, fig.height=4, fig.width=6}
# Load small demo data 
data("toyproblem")

# Unpack data
x <- toyproblem$x
dx <- x[2] - x[1]
y_obs <- toyproblem$y_obs
y_true <- toyproblem$y_true
dydx_true <- toyproblem$dydx_true


# Extimate derivative
dydx <- TVRegDiffR(
  data = y_obs,
  iter = 1e3,
  alph = 0.05,
  scale = "small",
  ep = 1e-6,
  dx = dx
  )
dydx <- dydx[-1]

# Prediction
y_pred <- y_obs[1] + cumsum(dydx*dx)

# Plot observed vs predicted
plot(x, y_obs, pch = 20, col = "#57575F", ylab = "y")
lines(x, y_pred, col = "#CA3542", lty = 2, lwd = 2)
lines(x, y_true, col = "#27647B", lty = 1, lwd = 2)
legend("bottomleft", 
       legend = c("Predicted", "True"),
       col = c("#CA3542", "#27647B"),
       lty = 2:1,
       cex = 0.8,
       lwd = 2)

# Plot derivative
plot(x, dydx, type = "l", ylim = c(-4, 4), col = "#CA3542", 
     ylab = expression(dy/dx), lty = 2, lwd = 2)
lines(x, dydx_true, col = "#27647B", lty = 1, lwd = 2)
legend("topleft",
       legend = c("Predicted", "True"),
       col = c("#CA3542", "#27647B"),
       lty = 2:1,
       cex = 0.8,
       lwd = 2)

# Simple finite difference
plot(x[1:(length(x) - 1)], diff(y_obs) / diff(x), type = "l", lty = 2, lwd = 2,
     xlab = "x", ylab = expression(dy/dx), col = "#CA3542")
lines(x, dydx_true, col = "#27647B", lty = 1, lwd = 2)
legend("topleft",
       legend = c("Finite Difference", "True"),
       col = c("#CA3542", "#27647B"),
       lty = 2:1,
       cex = 0.8,
       lwd = 2)

```




